How Diverse is the MoMA: While Modern, Not Diverse in Its Artists and Directors

### 2023-04-19 outputs/paper/paper.qmd
- Added to draft of paper.qmd - Added to references.bib based on additional content in paper.qmd

# Set up workspace 
library(tidyverse)
library(tidyr)
library(knitr)
library(kableExtra)
library(rstanarm)
library(modelsummary)
```

@@

Diversity in museums is a topic that is increasingly being discussed and raised as a problem in museums. While the main goal of museums is to serve as cultural institutions and a place of learning, oftentimes the objects on display or the decision-making staff who work behind the scenes do not meet this mission of inclusion.

This paper will investigate the level of diversity in one such museum: the Museum of Modern Art (also known as MoMA). Founded in 1929, the MoMA, as its name suggests, houses a collection of contemporary art pieces. It was originally created to go against traditional art museums by housing only modern art, and has since grown to become an influential cultural institution. Using data made publicly available on the MoMA's GitHub account, this paper aims to evaluate how diverse the artists of the artworks on display are, as well as how diverse the directors and department heads who chose to display these artworks are, in terms of gender and nationality. Thus, the estimand is: how much of the artists and directors in the MoMA's history are/were not white men.

It was found that the MoMA's artists and directors were predominantly men. A simple linear regression model shows that time affects this proportion of men.

The paper's structure begins by explaining the data used, followed by a linear model, a discussion of the relationship between variables in the data and in the model, and a discussion on key findings.

# Data {#sec-data}
### Data Source
In this report, the 25795 observations from the ____ and 65 observations from the _______ are used to investigate the level of diversity among MoMA decision-making directors and displayed artists. This data was obtained from the MoMA's publicly available GitHub, which has several public repositories containing data regarding the museum [@MoMAGithub]. For this analysis, the MoMA GitHub repositories of ____ [@____] and ____ [@__] were used.

This report was created using the R statistical programming language [@R]. For the results and analysis of this report, all figures were created using the tidyverse package [@tidyverse]. Additionally, the tables were created using the packages knitr [@knitr] and kableExtra [@kableExtra], and the graphs were created using the packages tidyr [@tidyr] and ggplot2 [@ggplot2].

### Variables of Interest

Our report selects 7 variables for analysis. We focus on 6 particular factors that we predicted affects the amount of Internet use: age, sex, race, number of children, marital status, and happiness. In this paper, we investigate the influence of these various factors on the total number of hours of weekly Internet usage. The packages tidyverse [@tidyverse], haven [@haven], and reshape2 [@reshape2] were used to clean the dataset by grouping or converting the questionnaire responses into categorical or textual answers.

The variables, Number of Children and Weekly Internet Hours, are recorded as a numerical value. Weekly Internet Hours has a maximum value of 168, as there are only 168 hours in a week. Age is also recorded as a numerical value, but is cleaned to represent ages in groups of ten years. The other variables, Sex, Race, Marital Status, and Happiness Level, are all recorded as a textual option for respondents to choose from.

Investigating artists:

@@

# Bar graph showing nationality distribution of Artists in MoMA exhibits
cleaned_momaexhibit_data |>
  filter(ExhibitionRole == "Artist") |>
  ggplot(mapping = aes(x = Nationality)) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  geom_bar(position = "dodge2")+
  labs(x = "Nationality", y = "Count of Artists in MoMA Exhibits") 
```

@@

@@

# Bar graph showing gender distribution of Artists in MoMA exhibits
cleaned_momaexhibit_data |> 
  filter(ExhibitionRole == "Artist") |>
  count(Gender) |>
  ggplot(mapping = aes(x = Gender, y = n)) +
  geom_col() + 
  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  labs(x = "Gender", y = "Count of Artists in MoMA Exhibits") 
```

@@

Investigating directors and department heads:

@@

@@

# Model
From the exploratory data analysis above, it is clear that both the number of female MoMA directors and the number of female artists displayed in the MoMA is much less than the number of male counterparts in these positions. This gender relationship is worth further investigation using linear models.

@@

Here we briefly describe the three linear regression models used to investigate this relationship between gender and time. Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model Set-Up
The first linear regression model regarding the effect of time on the percentage of female MoMA directors is as follows:
$$\hat{y} = \beta_0 + \beta_1x_{year}$$

@@

* Percentage of female MoMA directors, the response variable, is used as a quantitative measure of gender diversity that is between 0 and 100

* Year, the explanatory variable, is a numeric variable between 1929 (the founding year of the MoMA) and 2023 (the year that this report was published) indicating a specific year in the history of the MoMA

@@

* Percentage of female MoMA artists, the response variable, is used as a quantitative measure of gender diversity that is between 0 and 100

* Year, the explanatory variable, is a numeric variable between 1929 (the founding year of the MoMA) and 1990 (the year that the MoMA public GitHub has data until) indicating a specific year in the history of the MoMA

@@

* Percentage of female MoMA artists, the response variable, is used as a quantitative measure of gender diversity that is between 0 and 100

* Percentage of female MoMA directors, an explanatory variable, is a numeric variable between 0 and 100 indicating the gender distribution of the directors who had the decision-making power to choose the artists to be displayed

* Year, an explanatory variable, is a numeric variable between 1929 (the founding year of the MoMA) and 2023 (the year that this report was published) indicating a specific year in the history of the MoMA

All of the models are run in R [@citeR] using the `stats` package of @stats.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
#pp_check(first_model) +
#  theme_classic() +
#  theme(legend.position = "bottom")
#posterior_vs_prior(first_model) +
#  theme_minimal() +
#  scale_color_brewer(palette = "Set1") +
#  theme(legend.position = "bottom") +
#  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2
#plot(first_model, "trace")
#plot(first_model, "rhat")
```

>>>>

# Set up workspace 
library(tidyverse)
library(tidyr)
library(knitr)
library(kableExtra)
library(modelsummary)
```

@@

Diversity in cultural institutions is a topic that is increasingly being discussed and raised as a problem to be tackled, in museums in particular [@museumnext]. While the main goal of museums is to serve as cultural institutions and a place of learning, oftentimes the objects on display or the decision-making staff who work behind the scenes do not meet this mission of inclusion.

This paper will investigate the level of diversity in one such museum: the Museum of Modern Art (also known as MoMA). Founded in 1929, the MoMA, as its name suggests, houses a collection of contemporary art pieces. It was originally created to go against traditional art museums by housing only modern art, and has since grown to become an influential cultural institution [@MoMA]. Using data made publicly available on the MoMA's GitHub account, this paper aims to evaluate how diverse the artists of the artworks on display are, as well as how diverse the directors and department heads who chose to display these artworks are, in terms of gender and nationality. Thus, the estimand is: the percentage of artists, directors, and department heads in the MoMA's history who are / were not white men.

It was found that the MoMA's artists and directors were predominantly men. Simple linear regression models show that time affects this percentage of female artists, directors, and department heads in that the number of female artists has increased on average by ___ percentage points per year, while the number of female directors and department heads has increased on average by 0.25 percentage points per year. These findings are important because they provide numerical proof of the lack of gender diversity in the MoMA, both in the artists displayed and in the directors and department heads who choose the artists to display, which could motivate the MoMA to take active steps to better the gender representation within their institution.

This paper's structure will begin by explaining the MoMA's publicly made available data regarding its artists and staff that is the subject of analysis. It will then explore this data to understand overall trends in nationality and gender distribution in the MoMA. Thirdly, linear models will be used to further analyze the MoMA's gender distribution and its changes over time. Finally, key findings, limitations, and next steps will be discussed.

# Data {#sec-data}

## Data Source

In this report, the 32,537 observations from the MoMA's exhibition index dataset and 65 observations from the directors and department heads dataset are used to investigate the level of diversity among MoMA decision-making directors and displayed artists. This data was obtained from the MoMA's publicly available GitHub, which has several public repositories containing data regarding the museum [@MoMAGitHub]. For this analysis, the MoMA's "exhibitions" GitHub repository [@MoMAGitHubexhibitions] was used. The exhibition index dataset was created by a MoMA Archives team and contains information regarding the exhibitions featured in the MoMA from 1929 to 1989. The team is currently working on adding information on exhibits from 1990 onwards to the dataset. The directors and department heads dataset contains information regarding all of the directors of the entire museum and the department heads of curatorial departments within the museum from 1929 to the present.

This report was created using the R statistical programming language [@R]. For the results and analysis of this report, all figures were created using the tidyverse package [@tidyverse]. Additionally, the tables were created using the packages knitr [@knitr] and kableExtra [@kableExtra], the graphs were created using the packages tidyr [@tidyr] and ggplot2 [@ggplot2], and the models were displayed using the modelsummary package [@modelsummary].

## Variables of Interest

This report selects 7 variables for analysis from the MoMA exhibition index dataset and 6 variables from the directors and department heads. Variables regarding social background, specifically gender and nationality, were chosen in order to measure diversity within the MoMA.

@tbl-exhibits shows the 7 variables of the MoMA exhibition index dataset and the first 10 rows of it, where each row represents a different individual constituent in the exhibit. The Exhibition Title variable refers to the exhibit name. The Name variable indicates the artist. Nationality indicates the individual's nationality (e.g. "American") and is investigated further in @_____. Gender has two options of "Male" or "Female". Exhibition Start and Exhibition End indicate the time period that the exhibit was displayed at the MoMA for, and are the only numerical variables in this dataset.

```{r}
#| echo: false
#| eval: true
#| label: tbl-exhibits
#| tbl-cap: "First ten rows of the MoMA Exhibits dataset"
#| warning: false

cleaned_momaexhibit_data |> 
  slice(1:10) |> 
  knitr::kable(
    col.names = c("Exhibition Title", "Artist Name", "Nationality", "Gender", "Exhibition Start", "Exhibition End"),
    booktabs = TRUE,
    linesep = ""
  )
```

@tbl-directors shows the 6 variables of the MoMA directors and department heads dataset and the first 10 rows of it, where each row represents a different individual. The Department variable refers to the specific curatorial department within the MoMA that the individual is in charge of. As with the exhibition dataset, the Name variable indicates the person, Nationality indicates the individual's nationality (e.g. "American") and is investigated further in @_____, and Gender has two options of "Male" or "Female". Start and End indicate the time period that the individual held the position for, and are the only numerical variables in this dataset.

```{r}
#| echo: false
#| eval: true
#| label: tbl-directors
#| tbl-cap: "First ten rows of the MoMA Directors and Department Heads dataset"
#| warning: false

cleaned_momadirectors_data |> 
  slice(1:10) |> 
  knitr::kable(
    col.names = c("Department", "Name", "Start", "End", "Nationality", "Gender"),
    booktabs = TRUE,
    linesep = ""
  )
```

The data was cleaned and modified using the tidyverse package [@tidyverse].

## Data Visualization

The data was visualized as bar graphs in order to best understand the distribution of the diversity indicators in the data (i.e. the Gender and Nationality variables).

@fig-artistnationality displays the nationality distribution of the artists whose works have been displayed in exhibits in the MoMA. Aside from the large number of missing values (marked as NA at the rightmost bar), the most common nationality among is American, at 14,112 artists. The second highest nationality is French, at 3,747 artists - nearly a fifth of the American artists. The large number of American artists may be explained in part because the MoMA itself is American. It is surprising though that while the MoMA has presented 2,500+ exhibits throughout its history, \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_.

@@

# Bar graph showing nationality distribution of Artists in MoMA exhibits
cleaned_momaexhibit_data |>
  ggplot(mapping = aes(x = Nationality)) +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  geom_bar(position = "dodge2")+
  labs(x = "Nationality", y = "Number of Artists in MoMA Exhibits") 
```

@@

@fig-artistgender displays the gender distribution of the artists whose works have been displayed in exhibits in the MoMA. As the graph demonstrates, there is a much larger number of male artists who have been featured in the MoMA exhibits than female artists. With 22,518 male artists and 2,301 female artists, the male to female ratio is approximately 10:1, showing significant inequality in gender representation in MoMA artists. This may be partly explained by _____.

@@

# Bar graph showing gender distribution of Artists in MoMA exhibits
cleaned_momaexhibit_data |> 
  count(Gender) |>
  ggplot(mapping = aes(x = Gender, y = n)) +
  geom_col() + 
  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  labs(x = "Gender", y = "Count of Artists in MoMA Exhibits") 
```

@@

@fig-staffnationality displays the nationality distribution of all of the directors and department heads at the MoMA. Here too, like in the nationality distribution of artists, there is a majority of American nationality among the directors and department heads, at 50 Americans. Compared to the 3 Swiss, 3 British, 2 German, 1 French, and 1 Austrian directors and department heads, \_\_\_\_\_\_\_\_\_\_.

@@

@fig-staffgender displays the gender distribution of all of the directors and department heads at the MoMA. Here too, there are many more male directors and department heads compared to female department heads. With 53 male directors and department heads and 12 female directors and department heads, the male to female ratio is approximately 4:1, showing inequality in gender representation in MoMA leadership. This may be partly explained by \_\_\_\_\_\_\_\_\_\_.

@@

# Model

From the exploratory data analysis above, it is clear that both the number of female MoMA directors and the number of female artists displayed in the MoMA is much less than the number of male counterparts in these positions. This gender relationship is worth further investigation using linear models.

@@

Here we briefly describe the three linear regression models used to investigate this relationship between gender and time.

## Model Set-Up

The first linear regression model regarding the effect of time on the percentage of female MoMA directors is as follows: $$\hat{y} = \beta_0 + \beta_1x_{year}$$

@@

-   Percentage of female MoMA directors, the response variable, is used as a quantitative measure of gender diversity that is between 0 and 100

-   Year, the explanatory variable, is a numeric variable between 1929 (the founding year of the MoMA) and 2023 (the year that this report was published) indicating a specific year in the history of the MoMA

@@

-   Percentage of female MoMA artists, the response variable, is used as a quantitative measure of gender diversity that is between 0 and 100

-   Year, the explanatory variable, is a numeric variable between 1929 (the founding year of the MoMA) and 1990 (the year that the MoMA public GitHub has data until) indicating a specific year in the history of the MoMA

@@

-   Percentage of female MoMA artists, the response variable, is used as a quantitative measure of gender diversity that is between 0 and 100

-   Percentage of female MoMA directors, an explanatory variable, is a numeric variable between 0 and 100 indicating the gender distribution of the directors who had the decision-making power to choose the artists to be displayed

-   Year, an explanatory variable, is a numeric variable between 1929 (the founding year of the MoMA) and 2023 (the year that this report was published) indicating a specific year in the history of the MoMA

All of the models are run in R [@R] and displayed using modelsummary [@modelsummary].

### Model Justification

For the first two models regarding the effect of time on the percentage of female MoMA directors and on artists, time was chosen to be evaluated because there has been an increased push for gender equality in workplaces across the United States over the years. Because of this, a positive relationship between time in years and the percentage of female directors and female artists is expected for both the first linear regression model (regarding the effect of time on the percentage of female MoMA directors) and the second linear regression model (regarding the effect of time on the percentage of female MoMA artists).

Given these two models, it is worth investigating whether there is a relationship between the percentage of female directors and the percentage of female artists, to determine whether female directors are more likely to be aware of gender inclusion and thus display more female artists than their male counterparts. Because women are subjects of workplace gender discrimination, they are more likely to be aware of the need for inclusive gender representation, so a positive relationship between time in years and the percentage of female directors and female artists is expected.

@@
